---
title: "Maraca Plots - Animation"
author: "Kevin M. Anderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Maraca Plots - Basic Usage}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
library(dplyr)
library(ggplot2)
library(gganimate)
library(magick)
library(maraca)
```

The maraca package now supports built-in plot animations. This feature is useful for sprucing up presentations or highlighting the temporal nature of step-wise outcomes. 

In this vignette, we will show you how to easily animate a figure by calling the `animate_maraca` function. 


## Usage

### Basic Static Maraca Plot 

```{r, eval = TRUE, fig.width = 7, fig.height = 6}
# --- Load/Format Example Data --- 
data(hce_scenario_a, package = "maraca")
data          <- hce_scenario_a
column_names  <- c(outcome = "GROUP", arm = "TRTP",value = "AVAL0")
step_outcomes <- c("Outcome I", "Outcome II", "Outcome III", "Outcome IV")
last_outcome  <- "Continuous outcome"
arm_levels    <- c(active = "Active", control = "Control")

# --- Build Maraca Object --- 
mar <- maraca(
  data, step_outcomes, last_outcome, arm_levels, column_names, 
  fixed_followup_days = 3*365,
  compute_win_odds = TRUE
)

# --- Basic Static Maraca Plot --- 
p <- plot(mar, continuous_grid_spacing_x = 20)
p
```

### Animate Basic Maraca Plot

We create a GIF by simply calling the `animate_maraca` function, which takes all the same plotting parameters as the standard maraca `plot` function. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, continuous_grid_spacing_x = 20)
anim
```

Maraca color themes can be used for animations as well. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                        theme = 'color1',
                        continuous_grid_spacing_x = 20)
anim
```

### Controlling Animation Speed and Duration

There are two additional input parameters to control the speed and duration of the animation:

- __frames_per_step__: `int` or `list(int)`
  - This defines how many frames to generate for each x-axis __"step"__ of the maraca plot. A step is a single level of the outcome. For instance, the above plot has 5 total steps (Outcome I, Outcome II, Outcome III, Outcome IV, and Continuous).

- __gif_duration__: `int` or `float`
  - This defines how long the gif should be in seconds.

These examples show how to modify the overall gif duration.
```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
# 10 second GIF
slow_anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       gif_duration=10)
slow_anim
# 3 second GIF
fast_anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       gif_duration=3)
fast_anim
```

You can control the temporal granularity (i.e. the number of steps/frames) of the animation by setting the `frames_per_step` feature. 

Here we illustrate what a high and low frame temporal resolution looks like. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
# Low Temporal Granularity
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = 2,
                       gif_duration=5)
anim

```
```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
# High Temporal Granularity
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = 50,
                       gif_duration=5)
anim

```

You can also control the relative duration of each step being plotted. 

The `frames_per_step` parameter will accept a single integer if you want each segment/step to be the same length. 

Alternatively, it will accept a list of integers for more granular control. The length of `frames_per_step` must equal the number of `step_outcomes` + 1 (i.e. the last continuous outcome). 

In this example, we set the animation for `Outcome II` to have 4x the number of frames as the other steps.

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = c(10, 40, 10, 10, 10),
                       gif_duration=5)
anim
```

In this example, we have the animation pause for a long time on the continuous outcome (i.e. violins). 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = c(10, 10, 10, 10, 50),
                       gif_duration=5)
anim
```
\newpage


## Sequential Animation of Each Group/Arm

Thus far, we have animated the Active and Control arms simultaneously, but you have the ability to animate the two arms sequentially by setting the `arm_animation_order` parameter.  

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}

# This example will first animate the Control arm, then the Active Arm
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = 20,
                       arm_animation_order = c('Control', 'Active'),
                       gif_duration=8)
anim
```

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
# Or animate the "Active" arm first
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = 20,
                       arm_animation_order = c('Active', 'Control'),
                       gif_duration=8)
anim
```

You have similar control over the animation speed of each step.

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       frames_per_step = c(10,10,40,10,10),
                       arm_animation_order = c('Control', 'Active'),
                       gif_duration=8)
anim
```

## Animating Alternative Maraca Plots

The animation utility can handle any permutation of maraca plot you would like. 

For instance, we can animate a figure with boxplots instead of a violin plot. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20,
                       density_plot_type = "box",
                       frames_per_step=20, 
                       gif_duration=5)
anim
```

Sequential animation of each arm...
```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20,
                       density_plot_type = "box",
                       arm_animation_order = c('Control', 'Active'),
                       frames_per_step=20, 
                       gif_duration=8)
anim
```

...and scatter plots. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20, 
                       density_plot_type = "scatter", 
                       vline_type = "mean",
                       frames_per_step=20, 
                       gif_duration=5)
anim
```

...Sequential animation for the jitter/scatter continuous endpoint doesn't work well right now. There is some randomness to how the individual points are placed by ggplot2, which is not fixable with a simple `set.seed`. 

This is on the TO-DO list. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
                       continuous_grid_spacing_x = 20,
                       density_plot_type = "scatter", 
                       arm_animation_order = c('Control', 'Active'),
                       frames_per_step=20, 
                       gif_duration=8)
anim
```

We can also animate maraca plots with a binary final outcome. 

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
data("hce_scenario_a")

# --- Load/Format Example Binary Data --- 
bin_data <- hce_scenario_a
idx_cont <- bin_data$GROUP == "Continuous outcome"
bin_data[idx_cont,"GROUP"] <- "Binary outcome"
bin_data[idx_cont,"AVAL0"] <- bin_data[idx_cont,"AVAL0"] >= 0
bin_data[idx_cont,"AVAL"] <- bin_data[idx_cont,"AVAL0"] +
  bin_data[idx_cont,"GROUPN"]

column_names <- c(outcome = "GROUP", arm = "TRTP", value = "AVAL0")
step_outcomes <- c("Outcome I", "Outcome II",
                   "Outcome III", "Outcome IV")
last_outcome <- "Binary outcome"
arm_levels <- c(active = "Active", control = "Control")

# --- Create Maraca Object --- 
mar <- maraca(
  bin_data, step_outcomes, last_outcome,
  arm_levels, column_names,
  fixed_followup_days = 3*365,
  compute_win_odds = TRUE,
  last_type = "binary"
)

# --- Static Binary Plot --- 
plot(mar)
```

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
               frames_per_step=c(10,10,10,10,40), 
               gif_duration=5)
anim
```


```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
               frames_per_step=c(10,10,10,10,40),
               arm_animation_order = c('Control', 'Active'), 
               gif_duration=8)
anim
```


We can also animate outcomes/steps with a binary outcome

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
data("hce_scenario_a")

# --- Load/Format Example Binary Data --- 
bin_data2 <- hce_scenario_a
idx_bin <- bin_data2$GROUP %in% c("Outcome III", "Outcome IV")
bin_data2[idx_bin,"AVAL0"] <- bin_data2[idx_bin,"AVAL0"] >= 500
bin_data2[idx_bin,"AVAL"] <- bin_data2[idx_bin,"AVAL0"] +
  bin_data2[idx_bin,"GROUPN"]
bin_data2 <- bin_data2[bin_data2$AVAL0 != 0,]
head(bin_data2)

column_names <- c(outcome = "GROUP",arm = "TRTP",value = "AVAL0")
step_outcomes <- c("Outcome I", "Outcome II", "Outcome III", "Outcome IV")
last_outcome <- "Continuous outcome"
arm_levels <- c(active = "Active", control = "Control")

# --- Create Maraca Object --- 
mar <- maraca(
  bin_data2, step_outcomes, last_outcome,
  arm_levels, column_names,
  fixed_followup_days = 3*365,
  compute_win_odds = TRUE,
  step_types = c("tte","tte","binary","binary")
)
plot(mar)

```

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
               frames_per_step=c(20,20,20,5,20), 
               gif_duration=8)
anim
```

```{r, eval = TRUE, message=FALSE, warning=FALSE, fig.width = 7, fig.height = 6}
anim <- animate_maraca(mar, 
               frames_per_step=c(20,20,20,5,20), 
               arm_animation_order = c('Control', 'Active'),
               gif_duration=12)
anim
```


\newpage

## Saving the Animation

```{r, eval = TRUE, fig.width = 7, fig.height = 6}
anim_save(filename="~/my_animation.gif", animation=anim)
```
