---
title: "Maraca Plots - Basic Usage"
author: "Martin Karpefors, Samvel Gasparyan, Stefano Borini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Maraca Plots - Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
library(dplyr)
library(maraca)
```

# Introduction

Hierarchical composite endpoints (HCE) are complex endpoints combining events
of different clinical importance into an ordinal outcome that prioritize the
most severe event of a patient. Up to now, one of the difficulties in
interpreting HCEs has been the lack of proper tools for visualizing the
treatment effect captured by HCE. This package makes it possible to visualize
endpoints consisting of the combination of one to several time-to-event (TTE)
outcomes (like Death, Hospitalization for heart failure, …) with one continuous
outcome (like Kansas City Cardiomyopathy Questionnaire, KCCQ).

The maraca plot captures all important features of how the HCE measures the
treatment effect:

*1.* the relative proportions of time-to-event outcomes in the overall population,
*2.* the treatment effect on the combined and individual time-to-event outcomes, and
*3.* the treatment effect on the continuous component.


# Basic usage

Plotting with maraca requires to have the data in an appropriate format.
We provide a few scenarios in the package, named `hce_scenario_a.csv`,
`hce_scenario_b.csv`, `hce_scenario_c.csv` and `hce_scenario_d.csv`. These scenarios cover four cases: a) Treatment effect is driven by a combination of TTE outcomes and continuous outcome, b) No treatment effect on neither the TTE outcomes nor on the continuous outcome, c) A large treatment effect on time-to-event outcomes, but no effect on continuous outcome, and finally d) A small treatment effect on time-to-event outcomes, with a larger effect on continuous outcomes, respectively.
Other HCE scenarios can easily be simulated with the [hce package](https://github.com/Samve/hce). The hce package is also used for win odds calculations in the maraca package.

```{r maraca1, eval = TRUE}
file_path <- system.file("extdata", "hce_scenario_b.csv", package = "maraca")

data <- read.csv(file_path, stringsAsFactors = FALSE)
colnames(data)
```

The option `stringsAsFactors = FALSE` is required for R 3.x. It is the
default in R > 4.0 and can be omitted. The important point is that the
data frame must contain character strings, not factors.

The `data` data frame contains information on various columns. These columns
may have arbitrary names, so the `maraca` function allows you to specify
these names and how they map to the internal nomenclature using the
`column_names` parameter. This parameter is a named character vector, mapping
the standard names used by maraca with the column names in your dataset.

```{r maraca2, eval = TRUE}
column_names <- c(
    outcome = "GROUP",
    arm = "TRTP",
    value = "AVAL0"
)
```

the `outcome` column must be a character column containing the outcome
for each entry.

```{r maraca3, eval = TRUE}
unique(data[["GROUP"]])
```

The strings associated to each entry are arbitrary, so maraca
allows you to specify them in the `tte_outcomes` and `continuous_outcome`
parameters. Make sure to specify the TTE outcomes in the correct order,
starting from the most severe outcome to the least severe outcome. There can
only be one continuous outcome.

```{r maraca4, eval = TRUE}
tte_outcomes <- c(
  "Outcome I", "Outcome II", "Outcome III", "Outcome IV"
)
continuous_outcome <- "Continuous outcome"
```

the `arm` column must also be a character column describing to which arm
each row belongs.

```{r maraca5, eval = TRUE}
unique(data[["TRTP"]])
```

The strings associated to each entry are arbitrary, so maraca
allows you to specify them in the `arm_levels` parameter, as a named
vector of character strings. In this example, our file contains "Active"
and "Control" as identifiers, as seen in the output above, so we will specify

```{r maraca6, eval = TRUE}
arm_levels = c(active = "Active", control = "Control")
```

Finally, the `value` column must contain numerical values for the data.

All of the above can be aggregated in the call to create a maraca object from
the given dataset

```{r maraca7, eval = TRUE}
mar <- maraca(
  data, tte_outcomes, continuous_outcome, arm_levels, column_names, compute_win_odds = TRUE
)
```

The win odds can be calculated and will be annotated in the maraca plot if available- However, it is not calculated by default. Please set the `compute_win_odds = TRUE` to make the calculation. The resulting estimate, 95% confidence levels and p-value will be shown in the maraca object and can also be retrieved with

```{r, eval = TRUE}
mar$win_odds
```


# Plotting the resulting object

To create a maraca plot, simply pass the maraca object as the first argument to the `plot` function. In addition to the maraca object, the user can optionally supply input parameters for selecting the `continuous_grid_spacing`; transforming the x-axis using "sqrt" in the `trans` argument; selecting different representations of the continuous distribution density in the `density_plot_type`; and setting `vline_type` to "mean" or "median".

The following types of plots that are available: "default", "violin", "box",
"scatter". The default type is just a nested combination of violin and box
plot. Figure 1A-D in reference 1 below, can be recreated through applying the
maraca function with subsequent plotting on each of the scenario A-D datasets
provided in this package. As an example scenario A is shown in the following
plot. 

```{r maraca8, eval = TRUE, fig.width = 7, fig.height = 6}
plot(mar, continuous_grid_spacing_x = 20)
```

```{r maraca9, eval = TRUE, fig.width = 7, fig.height = 6}
plot(mar, continuous_grid_spacing_x = 20, density_plot_type = "box")
```

```{r maraca10, eval = TRUE, fig.width = 7, fig.height = 6}
plot(mar, continuous_grid_spacing_x = 20, density_plot_type = "scatter", vline_type = "mean")
```

The plot_maraca function returns a ggplot2 object. This function will not render the plot immediately so you will have to print() it.
Conveniently, you can customize your maraca plot with the comprehensive toolbox of ggplot2. For example:

```{r maraca11, eval = TRUE, fig.width = 7, fig.height = 6, message=FALSE, warning=FALSE}
p <- plot_maraca(mar, continuous_grid_spacing_x = 20, density_plot_type = "scatter", vline_type = "mean")
p +
  theme_bw() + 
  theme(axis.text.x.bottom = element_text(angle = 90, vjust = 0.5, hjust = 0.5))
```


Additionally, there are a few supporting plots that can be used to look closer
at the time-to-event outcomes.  `plot_tte_components` provides a one-row grid of
the TTE outcomes Kaplan-Meier plots.

```{r maraca12, eval = TRUE, fig.width = 7, fig.height = 6, message=FALSE, warning=FALSE}
plot_tte_components(mar)
```

`plot_tte_composite` plots the Kaplan-Meier plot for the composite of all specified events. The plot is also annotated with the hazard ratio (more info in `mar$survmod_complete`). 

```{r maraca13, eval = TRUE, fig.width = 7, fig.height = 6, message=FALSE, warning=FALSE}
plot_tte_composite(mar)
```


# References

Martin Karpefors, Daniel Lindholm and Samvel B. Gasparyan, "The maraca plot -- a novel visualization of hierarchical composite endpoints", submitted to Clinical Trials, August 2022.

